---
title: "Covid91 vaccine study Final2023"
author: "Karen Gomez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
Females=filter(FinalData,sex=="F")
LGBTQ=filter(FinalData,LGBTQ=="gay")
LGBTQ=filter(FinalData,LGBTQ=="straight")
DrugUser=filter(FinalDat,DrugUser=="no")
DrugUser=filter(FinalDat,DrugUser=="yes")
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$: The new preventative vaccine (COVID921) does not exhibit any significant differences in effectiveness across the variety of subpopulations tracked, including different genders (guys and gals), drug use, and sexual orientation.

$H_A$: The new preventative vaccine (COVID921) does exhibit significant differences in effectiveness across the variety of subpopulations tracked, including different genders (guys and gals), drug use, and sexual orientation.

## Methods

This dataset involves categorical variables, as both the vaccine's effectiveness and the subgroups can be categorized. The categorical nature of the variables allows us to use appropriate statistical methods for analysis. A barplot will be utilized to visually present the results, while a data table will summarize the numerical outcomes. Inferential statistics will be employed, specifically the chi-squared test and Fisher's exact test, to examine differences among the four subgroups.

## Males
### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```

In the second plot, which illustrates the percentage of males who took the vaccine compared to those who did not, it appears that the vaccinated males have a lower likelihood of contracting COVID-19. Conversely, in the first plot, the frequency of COVID-19 cases is slightly lower among vaccinated males compared to those who received the placebo drug. Overall, it is evident that the placebo group experienced the highest number of COVID-19 cases among males, indicating that the vaccine is effective in preventing COVID-19 in this subgroup.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment,data=FinalData)
rowPerc(table1)
colPerc(table1)
```

The table demonstrates that males who receive the vaccine are more likely to remain uninfected and not contract the virus.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
ftestM=fisher.test(table1)
ftestM
effecM=ftestM$estimate/(ftestM$estimate+1)*100
```

The inferential results indicate significant findings. The chi-squared test yielded a p-value of 0.00000000000012, which is smaller than the chosen significance level of 0.05, leading us to reject the null hypothesis. Similarly, Fisher's exact test provided a p-value of 0.00000000000019, further supporting the rejection of the null hypothesis. Additionally, the odds ratio shows that males who took the placebo are 1.7 times more likely to contract COVID-19 compared to those who received the vaccine. This finding highlights the substantial protective effect of the vaccine in preventing COVID-19 cases among males, underscoring its importance in reducing infection rates and promoting public health.

### Summary

For males, it is observed from the analysis and from the efficacy that the vaccine is effective in preventing them from getting the virus. For this reason, I recommend males to take the vaccine.

## Females
### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```

According to the plot, females who received the placebo exhibit a higher risk of contracting the virus compared to other groups.

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment,data=Females)
rowPerc(table1)
colPerc(table1)
```

The table indicates that females who took the placebo are at a higher likelihood of contracting the virus.

## LGBTQ



## Druggies


# Overall Results and Conclusions